Modular Verification Using Extended Finite Automata

### 2006-03-16 doc/EFA/efa.tex
*** empty log message ***

\section{The formal definition}
An Extended Finite Automaton (EFA) is an
augmentation of the regular automaton with guard and action
formulas associated to the transitions. The transitions in the EFA
are enabled if and only if a guard formula is true. Moreover, when
a transition in an EFA is taken, updating actions of a set of
variables may follow. The formal definition of an EFA is here
explained by extending the definition of the regular finite automaton.
Typically, system behaviours are modelled using deterministic finite-state automata.
Nondeterminism is used here to support the possibility of having different guards on transitions
that are triggered by common events. To avoid contradictions when synchronizing
automata, we do not allow this for the updating actions of variables
that are shared between automata.\\

@@

First, we add to $A$ a set $X=X_1\times ... \times X_n$, where
$X_k$, $k=1,...,n$, are finite nonempty sets. For instance, $X_k$
could be any finite subset of $\mathbb{Z}$ or the set of states in  some
other finite deterministic automaton. The finite set $X$ is the
domain of definition of an $n$-tuple of variables
$x=(x_1,...,x_n)$ with initial value $x_0=(x_{10},...,x_{n0})$. If
a variable corresponds to the set of states in an automaton, we
shall call it a state variable. The initial value of a state
variable must equal the initial state of its automaton and we do
not admit updating actions of a state variable. To tell apart the
ordinary variables and the state variables we use the subscripts
$v$ and $s$, respectively. Using this convention we divide the
domain of definition and the variables into parts, $X=X_v \times
X_s$ and $x=(x_v,x_s)$.

@@

The set of transitions in the extended automaton is
$T_e=\{(q,\sigma,x,p)\in Q \times \Sigma\times X\times Q:
\big(\delta(q,\sigma)=p \big)\wedge \big(g_{q
\sigma p}(x)=1\big)\}$. Notice that two transitions are
identical if and only if the events, the states and the values
of all variables are the same. Hence, a transition $(q,\sigma,p)$ in the
regular automaton $A$ is divided into as many transitions as the number
of elements in the set $\{x\in X:g_{q
\sigma p}(x)=1\}$. However, the
alphabets of the automata are the same and the language of the
extended automaton is subset of the language of the regular automaton:
\begin{eqnarray}
L(A_e)=\{s\in \Sigma^*:
\delta_e(q_i,s,x_0)!\}\subseteq \{s\in \Sigma^*:
\delta(q_i,s)!\}=L(A).
\end{eqnarray}
In some sense EFA's have small languages and a few number of
states to the expense of many transitions. This is an important observation
since it is the main reason why EFA's enable efficient modular algorithms to perform
verifications of supervisors.
\subsection{How to use the variables}
All logical systems can easily be represented by
EFA models. When using EFA's as a modelling tool it important to understand how to choose the variables.
A variable $x$ can represent many things, some examples  are: the states that an object can be in,
 the number of customers in a queue system, the different types of customers in a queue or the time it
takes for a process to execute different operations. In many applications the variables can be represented by
 integers or enumeration objects.

\subsubsection{Integer variables}

\subsubsection{Enumeration variables}

@@

An EFA can be translated into its corresponding FA.
If $A$ is the translation of $A_e$ we shall use the notation
$A=trans(A_e)$. When an EFA is "flatten out" the number of states
increases to $Q \times X_v $. The reason for this is that all the
actions are removed and translated into events in the regular FA. Notice
that when $A_e$ only has ordinary variables $X=X_v$. If $A_e$ also has
state variables the guard conditions on these variables are evaluated. When a variable
corresponds to the set of states
in an other automata this is not possible. Therefore its guard conditions are set to
TRUE when the translation is performed. To simplify the
algorithm we shall restrict us to quite simple action- and guard
functions. We shall not compare variables with each other in the guard functions.
The updating of the variables are assumed to be independent
from each other. This implies that they can be written as
tuples of functions of one variable i.e.
$a_{q \sigma p}(x)=(a^1_{q \sigma p}(x),...,a^n_{q \sigma p}(x))=(a^1_{q
\sigma p}(x_1),...,a^n_{q \sigma p}(x_n))$. We shall assume that the
guards are parsed and written in disjunctive normal form $g_{q
\sigma p}(x)=g_1(x)\vee ...\vee g_j(x)= (g^{1}_{1}(x_1)\wedge
...\wedge g^{n}_{1}(x_n) )\vee ...\vee (g^{1}_{j}(x_1)\wedge
...\wedge g^{n}_{j}(x_n) )$.

>>>>

\section{The formal definition}
An Extended Finite Automaton (EFA) is an
augmentation of the regular automaton with guard and action
formulas associated to the transitions. The transitions in the EFA
are enabled if and only if a guard formula is true. Moreover, when
a transition in an EFA is taken, updating actions of a set of
variables may follow. The formal definition of an EFA is here
explained by extending the definition of the regular finite automaton.
Typically, system behaviours are modelled using deterministic finite-state automata.
Nondeterminism is used here to support the possibility of having different guards on transitions
that are triggered by common events. To avoid contradictions when synchronizing
automata, we do not allow this for the updating actions of global variables
that are shared between automata.\\

@@

First, we add to $A$ a set $X=X_1\times ... \times X_n$, where
$X_k$, $k=1,...,n$, are finite nonempty sets. For instance, $X_k$
could be any finite subset of $\mathbb{Z}$, the set of states an automaton or
an input signal. The finite set $X$ is the
domain of definition of an $n$-tuple of variables
$x=(x_1,...,x_n)$ with initial value $x_0=(x_{10},...,x_{n0})$. If a variable is
an input signal to the system that $A$ describes or the value
of the variable represents the state that an automaton occupies, we
shall call it a state variable. The initial value of a state
variable must equal the initial state of its automaton and we do
not admit updating actions of a state variable. To tell apart the
ordinary variables and the state variables we use the subscripts
$v$ and $s$, respectively. Using this convention we divide the
domain of definition and the variables into parts, $X=X_v \times
X_s$ and $x=(x_v,x_s)$.

@@

The set of transitions in the extended automaton is
$T_e=\{(q,\sigma,x,p)\in Q \times \Sigma\times X\times Q:
\big(\delta(q,\sigma)=p \big)\wedge \big(g_{q
\sigma p}(x)=1\big)\}$. Notice that two transitions are
identical if and only if the events, the states and the values
of all variables are the same. Hence, a transition $(q,\sigma,p)$ in the
regular automaton $A$ is divided into as many transitions as the number
of elements in the set $\{x\in X:g_{q
\sigma p}(x)=1\}$. However, the
alphabets of the automata are the same and the language of the
extended automaton is subset of the language of the regular automaton:
\begin{eqnarray}
L(A_e)=\{s\in \Sigma^*:
\delta_e(q_i,s,x_0)!\}\subseteq \{s\in \Sigma^*:
\delta(q_i,s)!\}=L(A).
\end{eqnarray}
In some sense EFA's have small languages and a few number of
states to the expense of many transitions. This is an important observation
since it is the main reason why EFA's enable efficient modular algorithms to perform
verifications and to find supervisors.
\subsection{How to use the variables}
All logical systems can easily be represented by
EFA models. When using EFA's as a modelling tool it important to choose variables in a structured way.
A variable $x$ can represent many things, some examples  are: the states that an object can be in,
 the number of customers in a queue system, the different types of customers in a queue or the time it
takes for a process to execute different operations. In many applications the variables can be represented by
 integers or enumeration objects. One way to standardize model building with EFA is to each EFA divide the
 variables in input variables and output variables.

\subsubsection{Input variables}

\subsubsection{Output variables}

@@

An EFA can be translated into its corresponding FA.
If $A$ is the translation of $A_e$ we shall use the notation
$A=trans(A_e)$. When an EFA is "flatten out" the number of states
increases to $Q \times X_v $. The reason for this is that all the
actions are removed and translated into events in the regular FA. Notice
that when $A_e$ only has ordinary variables $X=X_v$. If $A_e$ also has
state variables the guard conditions on these variables can not be evaluated. Therefore its guard conditions are set to
TRUE when the translation is performed. To simplify the
algorithm we shall restrict us to quite simple action- and guard
functions. We shall not compare variables with each other in the guard functions.
The updating of the variables are assumed to be independent
from each other. This implies that they can be written as
tuples of functions of one variable i.e.
$a_{q \sigma p}(x)=(a^1_{q \sigma p}(x),...,a^n_{q \sigma p}(x))=(a^1_{q
\sigma p}(x_1),...,a^n_{q \sigma p}(x_n))$. We shall assume that the
guards are parsed and written in disjunctive normal form $g_{q
\sigma p}(x)=g_1(x)\vee ...\vee g_j(x)= (g^{1}_{1}(x_1)\wedge
...\wedge g^{n}_{1}(x_n) )\vee ...\vee (g^{1}_{j}(x_1)\wedge
...\wedge g^{n}_{j}(x_n) )$.
